---
title: "Rural Urban"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
  html_notebook:
    code_folding: hide
    float_toc:
      collapsed: yes
      smooth_scroll: yes
    toc: yes
  word_document:
    toc: yes
---


###DRAFT AIMS
**Aim 1**
**Aim 2**
**Aim 3**


###PRELIMINARY DATA

###DATA Sources

###Variables

###Methods
```{r}
# LIBRARIES ------------------------------------------
#install.packages("car")
library(tidyverse) # for various packages
library(haven) # for reading Stata file
library(labelled) # for getting variable labels
library(magrittr) # for extra %$% pipe
library(car) # for regression diagnostics
library(sandwich) # for heteroskedasticy-robust SEs
library(lmtest) # for coeftest (SEs for heteroskedasticity)
library(gridExtra) # for combining ggplots
library(tidyr)
library(knitr)

```

```{r datareadin}

ru <- read_sav("https://github.com/kijohnson/Rural-Urban-Cancer/blob/master/MO%20IL%20cancer.sav?raw=true")
View(ru)
```

##Create M:I ratio as a measure of survival. From
```{r}
ru$MtoI=ru$Mortality/ru$Incidence
```


###Mortality to Incidence Ratios with MO and IL data combined 

```{r sboth }

linearreg_ratio <-lm(formula = MtoI ~ as.factor(UIC_2013),data=ru)
summary(linearreg_ratio)

plot(ru$UIC_2013,ru$MtoI)
require('ggplot2')
ggplot(ru, aes(x=as.numeric(UIC_2013), y= MtoI)) +geom_smooth()#plotting the data for both 


```

###Survival for MO

```{r sMO}
ruMO <- subset(ru,State==0) #how to select a dataset with only certain characteristics #0 state is MO
ruIL <- subset(ru,State==1) 

linear_MO <-lm(formula = MtoI ~ as.factor(UIC_2013),data = ruMO)
summary(linear_MO)

plot(ruMO$UIC_2013,ruMO$MtoI)
require('ggplot2')
ggplot(ruMO, aes(x=as.numeric(ruMO$UIC_2013), y= ruMO$MtoI))+geom_line()
ggplot(ruMO, aes(x=as.numeric(ruMO$UIC_2013), y= ruMO$MtoI))+geom_smooth() #plotting the data for MO
```

###Survival for IL

```{r sIL}
linear_IL <-lm(formula = MtoI ~ as.factor(UIC_2013),data = ruIL)
summary(linear_MO)

plot(ruIL$UIC_2013,ruIL$MtoI)
require('ggplot2')
ggplot(ruIL, aes(x=as.numeric(ruIL$UIC_2013), y= ruIL$MtoI))+geom_line()
ggplot(ruIL, aes(x=as.numeric(ruIL$UIC_2013), y= ruIL$MtoI))+geom_smooth() #plotting the data for MO

```
##Group UIC_2013 codes into Metro (1,2) no-metro (3-12) and metro (1,2), micropolitan (3,5,8), and rural/noncore (4,6,7,9-12)
1 and 2 to be metro, 3,5,and 8 to be micropolitan, and then the rest are rural noncore.  
```{r}
ru$metro[ru$UIC_2013<=2]<-0
ru$metro[ru$UIC_2013>2]<-1

ru$metro2[ru$UIC_2013<=2]<-0
ru$metro2[ru$UIC_2013==3|ru$UIC_2013==5|ru$UIC_2013==8]<-1
ru$metro2[ru$UIC_2013==4|ru$UIC_2013==6|ru$UIC_2013==7|ru$UIC_2013>=9]<-2

ordered(ru$metro,levels = c(0, 1),labels = c("Metro", "Non-Metro"))
ordered(ru$metro2, levels = c(0,1,2), labels=c("Metro", "Micropolitan", "Rural/Noncare"))


```
#merge variable in with ruMO and ruIL datasets
```{r}
ruMO2<-merge(ruMO[c(1)], subset(ru, State==0), by="County_Name")
ruIL2<-merge(ruIL[c(1)], subset(ru, State==1), by="County_Name")

```

```{r}

linear_both <-lm(formula = MtoI ~ as.factor(ru$metro), data = ru)
summary(linear_both)

linear_MO <-lm(formula = MtoI ~ as.factor(ruMO2$metro), data = ruMO2)
summary(linear_MO)

linear_IL <-lm(formula = MtoI ~ as.factor(ruIL2$metro), data = ruIL2)
summary(linear_IL)
```

##
```{r}

linear_both <-lm(formula = MtoI ~ as.factor(ru$metro2), data = ru)
summary(linear_both)

linear_MO <-lm(formula = MtoI ~ as.factor(ruMO2$metro2), data = ruMO2)
summary(linear_MO)

linear_IL <-lm(formula = MtoI ~ as.factor(ruIL2$metro2), data = ruIL2)
summary(linear_IL)
```

